---
title: Supplementary Material for "Small sample evaluation of resampling-based inference
  for topological features of neuroimages"
author: Simon Vandekar, Kaidi Kang, Neil Woodward, Anna Huang, Maureen McHugo, Shawn
  Garbett, Jeremy Stephens, Russell T. Shinohara, Armin Schwartzman, and Jeffrey Blume
date: "2/7/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=10, cache=FALSE, dev = 'pdf')
path = Sys.getenv('PATH')
path = Sys.setenv('PATH'=paste(path, '/home/rstudio/.local/bin', sep=':'))
```



```{r simconfig, results='hide', eval=TRUE, results='asis'}
# install the latest versions of the packages to perform these analyses.
devtools::install_github('simonvandekar/pbj', ref='longitudinal')
devtools::install_github('statimagcoll/NIsim')
### LIBRARIES ###
library(RNifti) # Nifti I/O
library(parallel) # mclapply
library(splines) # ns
library(mmand) # spatial cluster functions
library(fslr) # imaging tools
library(progress) # progress bar
library(pbj) # pbj package
library(PDQutils) # edgeworth expansion stuff
library(NIsim) # simulation tools
library(papayar)

# number of cores for parallel things
ncores = 24


### LOAD IN DATA FROM DROPBOX ###
dbimagedir = '/media/disk2/pbj/data/rockland/neuroimaging'
#dbresimagedir = '~/pbj/data/abide/neuroimaging/cpac/alff_res'
#dbimagedir = '~/pbj/data/abide/neuroimaging/cpac/alff_cropped/'
#dbresimagedir = '~/pbj/data/abide/neuroimaging/cpac/alff_cropped_res/'
maskfile = '/usr/local/fsl/data/standard/MNI152_T1_2mm_brain_mask.nii.gz'
#maskfile = '/media/disk2/pbj/data/rockland/neuroimaging/MNI152_T1_2mm_brain_mask_slab.nii.gz'
overlapmaskfile = file.path(dbimagedir, 'overlap_mask_2mm.nii.gz')
#maskfile = '/media/disk2/pbj/data/rockland/neuroimaging/MNI152_T1_2mm_brain_mask_1vox.nii.gz'
dbdatafile = '/media/disk2/pbj/data/rockland/demographic/RocklandBehavioral.csv'
template = '/usr/local/fsl/data/standard/MNI152_T1_2mm_brain.nii.gz'

 # creates slab mask
mask = readNifti(maskfile)
slabs = round(dim(mask)[3]/2) + -5:5
sum(mask[,,slabs])
mask[,,-slabs] = 0
slabmaskfile = '/media/disk2/pbj/data/rockland/neuroimaging/MNI152_T1_2mm_brain_mask_slab.nii.gz'
writeNifti(mask, slabmaskfile)



# load in data and get directories
dat = read.csv(dbdatafile)
dat$dir = file.path(dbimagedir, dat$AnonymizedID, dat$IDandSession)
#dat$dir[which(!file.exists(dat$dir))]
dat = dat[ file.exists(dat$dir), ]
# some subjects have two image folders, this just grabs the first one.
dat$files = file.path(sapply(dat$dir, function(dir) list.files(dir, pattern='*', full.names=TRUE)[1]), 'GRAY_MNORM/mwp1t1.nii.gz')
dat$files2mm = gsub('.nii.gz', '_2mm.nii.gz', dat$files)
dat$files2mmsm4 = gsub('.nii.gz', '_sm4.nii.gz', dat$files2mm)


if(!all(file.exists(dat$files2mmsm4))){
  ## CREATE DOWNSAMPLED DATA
  invisible(mcmapply(flirt, infile=dat$files, outfile=dat$files2mm, opts = '-applyxfm', MoreArgs=list(reffile=template, retimg=FALSE), mc.cores = ncores ))
  ## SMOOTH DOWNSAMPLED DATA
  invisible(mcmapply(susan, file=dat$files2mm, outfile=dat$files2mmsm4, MoreArgs=list(sigma=4, dimg='3', n_usans='0'), mc.cores=ncores))
}
# view files
# papayar::papaya(c(template, dat$files2mm[1], dat$files2mmsm4[1]))
# set downsampled smoothed files as the main file
dat$files = dat$files2mmsm4

# rename some variables and do some data curation
dat$age = dat$CalculatedAge
dat$sex = dat$WhatIsYourSex
dat$race = ifelse(!dat$WhatIsYourRace %in% c('White', 'Black'), 'Other', dat$WhatIsYourRace)
dat$pdf = file.path(sapply(dat$dir, function(dir) list.files(dir, pattern='*', full.names=TRUE)[1]), 'PDF/catreport_t1.pdf')

dat$GMvol = sapply(dat$files, function(img){ sum(readNifti(img))} )

## CREATE MASK where 70% of subjects have image value larger than 0.1
# Was not used for the simulations... Need to rerun with this new mask.
# Method taken from this paper: https://www.sciencedirect.com/science/article/pii/S1053811908009786
if(!file.exists(overlapmaskfile)){
  imgs = readNifti(dat$files)
  mask = imgs[[1]]
  mask[,,] = 0
  imgs = apply(simplify2array(imgs), 1:3, function(v) mean(v>0.1))
  
  mask[imgs>0.7] = 1
  writeNifti(mask, file=overlapmaskfile)
}
#papayar::papaya(c(template, overlapmaskfile))
# replace mask file with GM mask made from the sample
tmpmask = paste0(tempfile(), '.nii.gz')
mask = readNifti(slabmaskfile) * readNifti(overlapmaskfile)
writeNifti(mask, tmpmask)
maskfile = tmpmask

### SIMULATION PARAMETERS ###
fakePolySimConfig = list(
  # vector of sample sizes to simulate
  ns = c(25, 50, 100, 200),
  # number of simulations to run
  nsim=100,
  # number of bootstraps
  nboot = 500,
  # cluster forming thresholds
  cfts.s = c(0.1, 0.25, 0.4),
  cfts.p = c(0.01, 0.001),
  
  # radius for spheres of signal.
  rs=c(8),
  
  #### MODEL FORMULAS FOR SIMULATIONS ####
  formres = paste0(" ~ sex + race + ns(age, df=10)" ),
  # need age_at_scan in both models for testing nonlinear functions
  form = paste0(" ~ sex + race + age + fake_covariate1 + scale(fake_covariate1^2) + scale(fake_covariate1^3)" ),
  formred = paste0(" ~ sex + race + age + fake_covariate1"),
  #  weights for each subject. Can be a character vector
  W = NULL,
  # where to put residuals
  resdir = 'resid',
  # where to output results
  simdir = '/media/disk2/temp',
  dat = dat,
  mask = maskfile,
  output = '/media/disk2/pbj/pbj_ftest/df2_polynomial.rdata',
  ncores = 10,
  method='bootstrap',
  syntheticSqrt= '/media/disk2/pbj/pbj_ftest/sqrtMat.rds'
)
# use betas = 0 for global null
# parameters = betas * sd(y)/sd(x).
fakePolySimConfig$betas = rep(0, length(fakePolySimConfig$rs))

# GROUP IS 1 DOF
sexSimConfig = fakePolySimConfig
sexSimConfig$form = paste0(" ~ sex + race + age" )
sexSimConfig$formred = paste0(" ~ race + age")
sexSimConfig$output = '/media/disk2/pbj/pbj_ftest/df1_sex.rdata'

# FAKE POLY IS 2 DOF (ABOVE)
#randomPolySimConfig = fakePolySimConfig
#randomPolySimConfig$form = paste0(" ~ sex + age + fake_covariate2 + scale(fake_covariate2^2) + scale(fake_covariate2^3)" )
#randomPolySimConfig$formred = paste0(" ~ sex + func_mean_fd + age_at_scan + fake_covariate2")
#randomPolySimConfig$output = '~/pbj/pbj_ftest/df2_polynomial_randomX.rdata'

# FAKE GROUP IS 3 DOF
fakeGroupSimConfig = fakePolySimConfig
fakeGroupSimConfig$form = paste0(" ~ sex + race + age + fake_group" )
fakeGroupSimConfig$formred = paste0(" ~ sex + race + age" )
fakeGroupSimConfig$output = '/media/disk2/pbj/pbj_ftest/df3_fakegroup.rdata'

# 3 DOF
ageSplineSimConfig = fakePolySimConfig
ageSplineSimConfig$form = paste0(" ~ sex + race + ns(age, df=4)" )
ageSplineSimConfig$formred = paste0(" ~ sex + race + age" )
ageSplineSimConfig$output = '/media/disk2/pbj/pbj_ftest/df4_agespline.rdata'

dat = dat[!apply(is.na(dat[,c('sex', 'age', 'race', 'GMvol') ]), 1, any ), ]

knitr::kable(as.data.frame(table1::table1(~ sex + age + race + GMvol, data=dat)), format = 'latex')
```

# Simulation functions
```{r simulationFunctions, eval=TRUE}
# simfunc should contain a data argument, which is defined within runSim
# Other arguments are identical across simulation runs.
simFunc = function(lmfull, lmred, mask, data, nboot, cfts.s=NULL, cfts.p=NULL, sim){
  # generate fake covariates
  data$fake_group = factor(ceiling(ppoints(nrow(data))*4 ) )
  data$fake_covariate1 = ppoints(nrow(data))
  data$fake_covariate2 = rnorm(nrow(data))

  # t transform, robust, estimate covariance
  #robustStatmap = lmPBJ(data$images, form=lmfull, formred=lmred, mask=mask, data=data, transform = 'none', HC3 = FALSE )
  # methods to try to make stats map more normal
  invisible(capture.output(HC3RobustStatmap <- lmPBJ(data$images, form=lmfull, formred=lmred, mask=mask, data=data, transform = 't', HC3 = TRUE )))
  # t transform, classical, estimate covariance
  invisible(capture.output(tStatmap <- lmPBJ(data$images, form=lmfull, formred=lmred, mask=mask, data=data, transform = 't', robust=FALSE, HC3=TRUE)))
  
  RobustpTFCE = ptfce.statmap(HC3RobustStatmap)
  TpTFCE = ptfce.statmap(tStatmap)
  

  
  statmaps = c('tStatmap','HC3RobustStatmap') # 'robustStatmap')#, ) 
  out = list()
  
  # if both are passed, only p-value thresholding is performed
  if(!is.null(cfts.p)){
    thrs = qchisq(cfts.p, df = HC3RobustStatmap$sqrtSigma$df, lower.tail = FALSE)
  } else if(!is.null(cfts.s)){
    thrs = (cfts.s^2*HC3RobustStatmap$sqrtSigma$n) + tRobustStatmap$sqrtSigma$df
  } else{
    stop('cfts.p or cfts.s must be specified.')
  }
  # Apply each of the sampling methods
  for(statmapname in statmaps){
    ### BOOTSTRAP METHODS
    statmap = get(statmapname)
    # normal bootstrap
    #pbjNorm = getBoots(pbjSEI(statmap, nboot = nboot, cfts.s = cfts))
    if(statmapname %in% c('HC3RobustStatmap')){
      pbjRadT <- pbjInference(statmap, nboot = nboot, method='wild', cft = thrs, mask=statmap$mask, runMode='bootstrap', max=TRUE, CMI=TRUE)$pbj
      pbjRadT$obsStat$tfce = c(maxZ = max(RobustpTFCE$Z), p05Thr = RobustpTFCE$fwer0.05.Z)
      #invisible(capture.output(pbjRadInd <- pbjInference(statmap, nboot = nboot, rboot = function(n){ (2*rbinom(n, size=1, prob=0.5)-1)}, method='independence', statistic=simStats, thr = thrs, mask=statmap$mask)))
      invisible(capture.output(pbjNormT <- pbjInference(statmap, nboot = nboot, rboot = function(n){ rnorm(n)}, method='wild', cft = thrs, mask=statmap$mask, runMode='bootstrap', max=TRUE, CMI=TRUE)$pbj))
      pbjNormT$obsStat$tfce = c(maxZ = max(RobustpTFCE$Z), p05Thr = RobustpTFCE$fwer0.05.Z)
      invisible(capture.output(pbjPermT <- pbjInference(statmap, nboot = nboot, method='permutation', cft = thrs, mask=statmap$mask, runMode='bootstrap', max=TRUE, CMI=TRUE)$pbj))
      pbjPermT$obsStat$tfce = c(maxZ = max(RobustpTFCE$Z), p05Thr = RobustpTFCE$fwer0.05.Z)
      #pbjNonparametric = pbjInference(statmap, nboot = nboot, method='nonparametric', thr = thrs, mask=statmap$mask, statistic=simStats)
    }
    
     if(statmapname %in% c('tStatmap')){
      invisible(capture.output(pbjRadT <- pbjInference(statmap, nboot = nboot, method='wild', cft = thrs, mask=statmap$mask, runMode='bootstrap', max=TRUE, CMI=TRUE)$pbj ))
       pbjRadT$obsStat$tfce = c(maxZ = max(TpTFCE$Z), p05Thr = TpTFCE$fwer0.05.Z)
      invisible(capture.output(pbjNormT <- pbjInference(statmap, nboot = nboot, rboot = function(n){ rnorm(n)}, method='wild', cft = thrs, mask=statmap$mask, runMode='bootstrap', max=TRUE, CMI=TRUE)$pbj ))
      pbjNormT$obsStat$tfce = c(maxZ = max(TpTFCE$Z), p05Thr = TpTFCE$fwer0.05.Z)
      invisible(capture.output(pbjPermT <- pbjInference(statmap, nboot = nboot, method='permutation', cft = thrs, mask=statmap$mask, runMode='bootstrap', max=TRUE, CMI=TRUE)$pbj ))
      pbjPermT$obsStat$tfce = c(maxZ = max(TpTFCE$Z), p05Thr = TpTFCE$fwer0.05.Z)
      #pbjNonparametric = pbjInference(statmap, nboot = nboot, method='nonparametric', statistic=simStats, thr = thrs, mask=statmap$mask)
    }
    # collect output
    PBJnames = grep('^pbj', ls(), value=TRUE)
    allnames = paste(statmapname, PBJnames, sep='_')
    out[allnames] = lapply(PBJnames, get, pos = environment())
    rm(list=PBJnames)

  }
  gc()
  return(out)
}

#debug(pbjInference)
#simConfig = get("fakeGroupSimConfig")
#simdirs = simSetup(simConfig$dat$files, data=simConfig$dat, outdir=simConfig$simdir, nsim=simConfig$nsim, ns=simConfig$ns, mask=simConfig$mask, rs=simConfig$rs, betas=simConfig$betas )
#simtime = system.time(test <- simFunc(simConfig$form, simConfig$formred, simConfig$mask, readRDS("/media/disk2/temp/sim1/n100/data.rds"), sim = 1, 2, cfts.p = simConfig$cfts.p))
#stop('not an error.')
sims = grep('SimConfig', ls(), value=TRUE)
```





```{r runSims}
for(sim in sims[-c(1,2)]){
  # get simulation configuration for this simulation
  simConfig = get(sim) 
  ### SETUP THE SIMULATION ANALYSIS ###
  # subsets dataset to all people who have the variables
  simConfig$dat = simConfig$dat[apply(!is.na(simConfig$dat[ ,c(all.vars(as.formula(simConfig$formres)), simConfig$W)]), 1, all), ]
  # Create residualized images
  if(class(simConfig$formres)=='formula' | is.character(simConfig$formres)){
    simConfig$dat$rfiles = file.path(dirname(simConfig$dat$files), '..', simConfig$resdir, basename(simConfig$dat$files))
    # remake these files each time
    unlink(simConfig$dat$rfiles)
    if(!all(file.exists(simConfig$dat$rfiles))){
      residualizeImages(files=simConfig$dat$files, dat=simConfig$dat, mask=simConfig$mask, form=as.formula(simConfig$formres), outfiles=simConfig$dat$rfiles, mc.cores=simConfig$ncores, outrds=simConfig$syntheticSqrt)
    }
    simConfig$dat$files = simConfig$dat$rfiles
    # clean up. May not be necessary
    gc()
  }
  
  simdirs = simSetup(simConfig$dat$files, data=simConfig$dat, outdir=simConfig$simdir, nsim=simConfig$nsim, ns=simConfig$ns, mask=simConfig$mask, rs=simConfig$rs, betas=simConfig$betas )
  gc()
  
  
  #time = system.time(test <- simFunc(simConfig$form, simConfig$formred, simConfig$mask, readRDS(file.path(simdirs$simdir[10], 'data.rds')), simConfig$nboot, simConfig$cfts.s) )
  
  # mix this up so that large sample simulations aren't all dropped on one "thread".
  simdirs = simdirs[sample(1:nrow(simdirs)),]
  
  # check failed simulations
  if(file.exists(simConfig$output)){
    load(simConfig$output)
    if(length(dim(results))<=1){
      failedSims = sapply(results, length)<=1
      failedSims = names(failedSims)[failedSims]
      simSet = simdirs[ simdirs$simdir %in% failedSims,]
      
      rerunResults = runSim(simSet$simdir, method=simConfig$method,
                   simfunc = simFunc, mask = simConfig$mask, sims = simSet$sim,
                   simfuncArgs = list(
                     lmfull= simConfig$form,
                     lmred = simConfig$formred,
                     mask = simConfig$mask, nboot=simConfig$nboot, cfts.p=simConfig$cfts.p), ncores = simConfig$ncores, syntheticSqrt = simConfig$syntheticSqrt)
      rerunResults = unlist(apply(rerunResults, 2, list), recursive=FALSE)
      results[failedSims] = rerunResults
    }
  } else {
    failedSims = NULL
    results = runSim(simdirs$simdir, method=simConfig$method,
                   simfunc = simFunc, mask = simConfig$mask, sims = simdirs$sim,
                   simfuncArgs = list(
                     lmfull= simConfig$form,
                     lmred = simConfig$formred,
                     mask = simConfig$mask, nboot=simConfig$nboot, cfts.p=simConfig$cfts.p), ncores = simConfig$ncores, syntheticSqrt = simConfig$syntheticSqrt)
  }
  
  
  dir.create(dirname(simConfig$output), showWarnings = FALSE, recursive = TRUE)
  # clean up files
  save.image(file=simConfig$output)
  unlink(list.files(tempdir(), full.names = TRUE))
  gc()
  unlink(simdirs)
}
stop('not an error. Finished simulations.')
```





```{r, eval=TRUE}

# lis A list of numeric vectors. Length of list is equal to number of bootstraps or number of simulations. Each element of the list is a vector of cluster sizes or local maxima within that simulation.
# probs probabilities for quantiles 
quantileMarg = function(lis, probs=c(0.95, 0.99), na.rm=TRUE){ 
  nclusts = sapply(lis, function(y) sum(!is.na(y)))
  if(length(nclusts)!=0){
  ans = Hmisc::wtd.quantile(unlist(lis), weights = rep(1/nclusts, nclusts), probs=probs, na.rm=na.rm)
  } else {
    ans = quantile(NA, probs=probs, na.rm=TRUE)
  }
  ans
}


KL = function(lis1, lis2, np=100, na.rm=TRUE){ 
  nclusts1 = sapply(lis1, function(y) sum(!is.na(y)))
  nclusts2 = sapply(lis2, function(y) sum(!is.na(y)))
  rang = range(unlist(c(lis1, lis2)), na.rm=TRUE, finite=TRUE)
  ev = locfit::lfgrid(np, rang[1], rang[2])
  if(length(nclusts1)!=0){
    l1NAs = is.na(unlist(lis1)) | is.infinite(unlist(lis1))
    l2NAs = is.na(unlist(lis2)) | is.infinite(unlist(lis2))
    ans = locfit::density.lf(unlist(lis1)[!l1NAs], n = 100, weights=rep(1/nclusts1, nclusts1)[!l1NAs], ev = ev)
    ans2 = locfit::density.lf(unlist(lis2)[!l2NAs], n = 100, weights=rep(1/nclusts2, nclusts2)[!l2NAs], ev = ev)
    zeros = ans$y!=0 & ans2$y!=0
    # trapezoidal rule to approximate integral
    ans2$y[zeros] = ans2$y[zeros]/sum(ans2$y[zeros])
    ans$y[zeros] = ans$y[zeros]/sum(ans$y[zeros])
    ans = sum(ans2$y[zeros]*log(ans2$y[zeros]/ans$y[zeros]))
  } else {
    ans = NA
  }
  ans
}

cdfMarg = function(lis, tstat, na.rm=TRUE){ 
  nclusts = sapply(lis, function(y) sum(!is.na(y)))
  if(length(tstat)==0) tstat = 0
  if(sum(nclusts)==0){
    ans = rep(NA, length(tstat))
    } else {
     # if(length(rep(rep(1/nclusts, nclusts))) != length(unlist(lis))){
     #   browser()
     # }
      #tryCatch({
  ans = colSums(matrix(rep(rep(1/nclusts, nclusts), length(tstat)), ncol=length(tstat)) * outer(unlist(lis), tstat, '>=' ), na.rm=TRUE)/length(lis)
  # },
  # error = function(e){ recover()})
    }
  ans
}



# Creates plots for figures:
# First loop through methods to get plot data.
# Then plot type 1 error rates for all the local methods.
# Then plot type 1 error rates for all the global methods.
# Then plot KL divergence for all the local methods.
plotData = function(rdata, alpha=0.1, stats=NULL){
  load(rdata)
  
  plotDFs = list()
  qDFs = list()
  klDFs = list()
  tfce = list()
  # This is what should happen if nothing fails. If things fail it will still try to plot the results
  if(length(dim(results))>1){
    results = unlist(apply(results, 2, list), recursive=FALSE)
  }
  simdirs$results = results# lapply(results, simplify2array)
  methods = names(simdirs$results[[2]])
  if(is.null(stats)){
    stats = c("Maxima", paste('Extent; cft =', simConfig$cfts.p), paste('Mass; cft =', simConfig$cfts.p) )
  }
  
  for(method in methods){
    methodname = paste(ifelse(grepl('^tStatmap', method), 'Parametric', 'Robust'), ifelse(grepl('Norm', method), 'Normal bootstrap', ifelse(grepl('Rad', method), 'Rademacher bootstrap', 'Permutation') ) , sep="; ")
    obsStat = do.call(rbind, lapply(simdirs$results, function(y) if(is.null(y)) NA else y[[method]][['obsStat']] ) )
    colnames(obsStat) = stats
    # These colnames were sample size dependent
    maximas = as.data.frame(matrix(unlist(apply(obsStat, 2, function(x) lapply(x, max))), nrow=nrow(obsStat)))
    names(maximas) = paste('Global', stats)
    gtfce = which(names(maximas)=='Global pTFCE')
    if(length(gtfce)==1){
      maximas[, 'Global pTFCE'] = sapply(obsStat[, 'pTFCE'], function(x) x[1])^2
    }
    # add global maximums to list of statistics
    obsStat = cbind(obsStat, maximas)
    
    
    # Arrange each bootstrap like the obsStat setup
    simdirs$boots = lapply(simdirs$results, function(y) do.call(rbind, lapply(y[[method]][['boots']], function(z0) simplify2array(z0) ) ) )
    simdirs$boots = lapply(simdirs$boots, function(boot){
      # as written, TFCE needs to be the last statistic
      if(stats[length(stats)]=='pTFCE'){
        boot = cbind(boot, sapply(boot[, 'maxima'], max))
      }
      colnames(boot) = stats
      maximas = as.data.frame(matrix(unlist(apply(boot, 2, function(x) lapply(x, function(y) max(c(y,0)))) ), nrow=nrow(boot)))
      names(maximas) = paste('Global', stats)
      boot = cbind(boot, maximas)
    })
    #length.out=pmin(simConfig$nsim, simConfig$nboot)
    
    # for(cftInd in 1:ncol(obsStat)){
    #   #cat(cftInd, '\n')
    #   colname = colnames(obsStat)[cftInd]
    #   kldf = do.call(rbind, lapply(split(cbind(simdirs, obsStat[colname]), simdirs$n), function(df){
    #     #cat(df$n[1], '\n')
    #     KLs = sapply(df$boots, function(x){KL(x[,colname], df[[colname]], np = 100) } )
    #     ans = data.frame(n=df$n[1], KL=KLs)
    #     ans
    #   }) )
    #   #cat('\n')
    #   colnames(kldf)[2] = method
    #   if(method == methods[1]){
    #     klDFs[[colname]] = kldf
    #   } else {
    #     klDFs[[colname]] = cbind(klDFs[[colname]], kldf[,2])
    #   }
    # }
    
    tfceOut = sapply(split(obsStat$pTFCE, simdirs$n), function(x) mean(sapply(x, function(y){y[1]>y[2]})) )
    tfceOut = data.frame(n=names(tfceOut), 'Error'=tfceOut)
    tfce[[method]] = tfceOut
 
    xaxlab = c(0.75, 0.9, 0.95, 0.99)
    for(cftInd in 1:ncol(obsStat)){
      colname = colnames(obsStat)[cftInd]
      adjustMethod = if(grepl('Global', colname)) 'none' else 'BH'
      
      cat(colname)
      plotData = do.call(rbind, lapply(split(cbind(simdirs, obsStat[colname]), simdirs$n), function(df){
        
        pvalues = mapply(cdfMarg, lapply(df$boots, function(x) x[,colname]), df[,colname] )
        minPvalues = sapply(pvalues, function(pvalue) min(p.adjust(pvalue, method=adjustMethod), na.rm=TRUE))
        #ylims = range(quantiles, na.rm=TRUE )
        x = 1-xaxlab
        y = colMeans(outer(minPvalues, x, '<='), na.rm=TRUE)
        out = data.frame(x=x, y=y, n=df$n[1])
      }) )
      names(plotData)[2] = method
      # put the output into the variable colname
      if(method == methods[1]){
        plotDFs[[colname]] = plotData
      } else {
        plotDFs[[colname]] = merge(plotDFs[[colname]], plotData)
      }
    }
    
    
    ## For the QQ-plot
    xaxlab = c(0.75, 0.9, 0.95, 0.99)
    for(cftInd in 1:ncol(obsStat)){
      colname = colnames(obsStat)[cftInd]
      
      xlims = range(unlist(lapply(split(obsStat[[colname]], simdirs$n), quantileMarg, probs=xaxlab) ))
      
      plotData = lapply(split(cbind(simdirs, obsStat[colname]), simdirs$n), function(df){
        quantiles = sapply(df$boots, function(x) quantileMarg(x[,colname], probs=xaxlab) )
        #ylims = range(quantiles, na.rm=TRUE )
        
        x = quantileMarg(df[,colname], probs=xaxlab)
        # first column is observed quantiles across the simulations
        quantiles = cbind(x, quantiles)
        return(quantiles)
        #plot(x, ylim=ylims, xlim=xlims, type='n', ylab='', xlab='', main=paste0('n = ', df$n[1], '; ',  colname))
        #axis(side=1, at=xaxt, labels=xaxlab)
        #abline(v=xaxt, col='orange', lty=2)
        #for(ind in 1:simConfig$nboot){
        #  points(x, quantiles[,ind], type='l')
        #}
        #abline(a=0,b=1, col='blue')
      })
      ns = as.numeric(names(plotData))
      plotData = as.data.frame(do.call(rbind, plotData))
      plotData$n = rep(ns, each=length(xaxlab))
      plotData$method = method
      if(method == methods[1]){
        qDFs[[colname]] = plotData
      } else {
        qDFs[[colname]] = rbind(qDFs[[colname]], plotData)
      }
    }
    
 
  }
  #klDFs = lapply(klDFs, function(x){names(x)[-1] = methods; x})
  return(list(error=plotDFs, qq=qDFs, tfce=tfce)) #KL=klDFs, 
}


errorPlots = function(plotData, nsim, tStat=TRUE, global=FALSE, alpha=0.1){
  
  # graphical parameters
  cex=1.5
  par(mgp=c(1.8,.7,0), lwd=1.5, lend=2, cex.lab=0.8*cex, cex.axis=0.8*cex, cex.main=0.6*cex, mfrow=c(1,1), mar=c(1.7,1.7,1.8,.5), bty='l', oma=c(2,2,2,0))
  brewcols =  rep(RColorBrewer::brewer.pal(n=length(methods)/2, name='Dark2'), 2)
  
  lis = plotData[['error']]
  stats = grep('Global', names(lis), invert=!global, value=TRUE)
  if(!global){ # don't plot tfce for non global statistics
    stats = grep('TFCE', stats, value=TRUE, invert=TRUE)
  }
  nplots = length(stats) * length(unique(lis[[stats[1]]]$n))
  layout(matrix(1:(nplots), nrow=length(stats), byrow=TRUE) )
  # for each type of statistic, make a row of plots
  for(stat in stats){
    # for each sample size, make a plot
    lapply(split(lis[[stat]], lis[[stat]]$n), function(df){
           n = df$n[1]
           x = df$x
           methods = names(df)[ ! names(df) %in% c('x', 'n')]
           
           methodnames = paste(ifelse(grepl('tStatmap', methods), 'Parametric', 'Robust'), ifelse(grepl('Norm', methods), 'Normal bootstrap', ifelse(grepl('Rad', methods), 'Rademacher bootstrap', 'Permutation') ) , sep="; ")
           
           # T-statistics
           tmethodinds = grep('^tStatmap_', methods, invert=!tStat)
           rang = range(c(df[,methods[tmethodinds]], x))
           
           plot(x, df[,methods[tmethodinds[1] ] ], type='b', xlim=rang, ylim=rang, ylab='', xlab='', main=paste0('n = ', n, '; ',  stat), col=brewcols[1])
           for(ind in tmethodinds[-1]){
             points(x, df[,methods[ind]], type='b', col=brewcols[ind])
           }
           points(x, qbinom(alpha/2, nsim, x)/nsim, type='l', lty=2, col='gray')
           points(x, qbinom(1-alpha/2, nsim, x)/nsim, type='l', lty=2, col='gray')
           abline(a=0,b=1, col='gray')
           # hard-coded if statement
           if(stat==stats[1] & n==25){
           legend('bottomright', # Find suitable coordinates by trial and error
                  c('NB', 'Perm', 'RB'), fill=brewcols[1:3], bty='n', cex=1.2)
           }
    })
  }
  mtext(ifelse(tStat, 'Parametric', 'Robust'), outer=TRUE, font=2)
  mtext('Target type 1 error', outer=TRUE, side = 1)
  mtext('Actual type 1 error', outer=TRUE, side=2)
}



klPlots = function(plotData, nsim, tStat=TRUE, global=FALSE, alpha=0.1){
  
  # graphical parameters
  cex=1.5
  par(mgp=c(1.8,.7,0), lwd=1.5, lend=2, cex.lab=0.8*cex, cex.axis=0.8*cex, cex.main=0.6*cex, mfrow=c(1,1), mar=c(1.7,1.7,1.8,.5), bty='l', oma=c(2,2,2,0))
  
  lis = plotData[['KL']]
  stats = grep('Global', names(lis), invert=!global, value=TRUE)
  nplots = length(stats) * length(unique(lis[[stats[1]]]$n))
  layout(matrix(1:(nplots), nrow=length(stats), byrow=TRUE) )
  # for each type of statistic, make a row of plots
  for(stat in stats){
    # for each sample size, make a plot
    lapply(split(lis[[stat]], lis[[stat]]$n), function(df){
           n = df$n[1]
           methods = names(df)[ ! names(df) %in% c('x', 'n')]
           
           brewcols =  rep(RColorBrewer::brewer.pal(n=length(methods)/2, name='Dark2'), 2)
           methodnames = paste(ifelse(grepl('tStatmap', methods), 'Parametric', 'Robust'), ifelse(grepl('Norm', methods), 'Normal bootstrap', ifelse(grepl('Rad', methods), 'Rademacher bootstrap', 'Permutation') ) , sep="; ")
           
           # T-statistics
           tmethodinds = grep('^tStatmap_', methods, invert=!tStat)
           #rang = range(c(df[,methods[tmethodinds]]))
           
           pm = as.matrix(df[,methods[tmethodinds]])
           pm[is.infinite(pm)] = NA
           vioplot::vioplot(pm, col = brewcols[tmethodinds],
        xlab = "", ylab = "", main=paste0('n = ', n, '; ',  stat), names=c('NB', 'Perm', 'RB'))
    })
  
  mtext(ifelse(tStat, 'Parametric', 'Robust'), outer=TRUE, font=2)
  mtext('Method', outer=TRUE, side = 1)
  mtext('KL Divergence', outer=TRUE, side=2)
  }
}



qqPlots = function(plotData, nsim, tStat=TRUE, global=FALSE, alpha=0.1){
  
  # graphical parameters
  cex=1.5
  par(mgp=c(1.8,.7,0), lwd=1.5, lend=2, cex.lab=0.8*cex, cex.axis=0.8*cex, cex.main=0.6*cex, mfrow=c(1,1), mar=c(1.7,1.7,1.8,.5), bty='l', oma=c(2,2,2,0))
  
  lis = plotData[['qq']]
  stats = grep('Global', names(lis), invert=!global, value=TRUE)
  if(!global){ # don't plot tfce for non global statistics
    stats = grep('TFCE', stats, value=TRUE, invert=TRUE)
  }
  nplots = length(stats) * length(unique(lis[[stats[1]]]$n))
  layout(matrix(1:(nplots), nrow=length(stats), byrow=TRUE) )
  # for each type of statistic, make a row of plots
  for(stat in stats){
          methods = unique(lis[[stat]]$method)
          methodnames = paste(ifelse(grepl('tStatmap', methods), 'Parametric', 'Robust'), ifelse(grepl('Norm', methods), 'Normal bootstrap', ifelse(grepl('Rad', methods), 'Rademacher bootstrap', 'Permutation') ) , sep="; ")
          brewcols =  rep(RColorBrewer::brewer.pal(n=length(methods)/2, name='Dark2'), 2)
          # T-statistics
          tmethodinds = grep('^tStatmap_', methods, invert=!tStat, value=TRUE)
          yrang = quantile(unlist(lis[[stat]][lis[[stat]]$method %in% tmethodinds,2:(nsim+1)]), probs = c(0.01, 0.99))
          xrang = range(lis[[stat]][lis[[stat]]$method %in% tmethodinds,1])
    # for each sample size, make a plot
    lapply(split(lis[[stat]], lis[[stat]]$n), function(df){
           n = df$n[1]
           x = df$x[df$method==tmethodinds[1]]
           plot(x, df[df$method==tmethodinds[1], 2], type='n', xlim=xrang, ylim=yrang, ylab='', xlab='', main=paste0('n = ', n, '; ',  stat))
           for(ind in 1:nsim){
             points(x, df[df$method==tmethodinds[1], ind+1], type='l', lwd=0.5, col=brewcols[1]) # , col=brewcols[ind]
             points(x, df[df$method==tmethodinds[2], ind+1], type='l', lwd=0.5, col=brewcols[2])
             points(x, df[df$method==tmethodinds[3], ind+1], type='l', lwd=0.5, col=brewcols[3])
           }
           #points(x, qbinom(alpha/2, nsim, x)/nsim, type='l', lty=2, col='gray')
           #points(x, qbinom(1-alpha/2, nsim, x)/nsim, type='l', lty=2, col='gray')
           abline(a=0,b=1, col='gray')
           # hard-coded if statement
           if(stat==stats[1] & n==25){
           legend('topleft', # Find suitable coordinates by trial and error
                  c('NB', 'Perm', 'RB'), fill=brewcols[1:3], bty='n', cex=1.2)
           }
    })
  }
  mtext(ifelse(tStat, 'Parametric', 'Robust'), outer=TRUE, font=2)
  mtext('Simulated quantiles', outer=TRUE, side = 1)
  mtext('Resampled quantiles', outer=TRUE, side=2)
}
```



## Simulation results

Details of the simulation analyses and evaluation metrics are given in Section 4 of the paper. This section presents results for all simulations for parametric and robust test statistics evaluating the marginal and global CDFs.

## Sex covariate

The effect of sex was removed from the imaging data using equation (13), and then bootstrap samples of the residuals were modeled as a function of sex and tested on one degree of freedom.
Type 1 error rates and QQ-plot are given below.


```{r, eval=FALSE, fig.width=8, fig.height=10, fig.cap="Actual versus target type 1 error rates for the inference procedures considered for testing the marginal distribution of each topological feature (TF) of the parametric test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
pd = plotData(sexSimConfig$output)
nsim = fakeGroupSimConfig$nsim
errorPlots(pd, nsim=nsim, tStat = TRUE, global=FALSE)

#qqPlots(pd, nsim=nsim, tStat=TRUE, global=FALSE)
```

```{r, eval=FALSE, fig.width=8, fig.height=10, fig.cap="Actual versus target type 1 error rates for the inference procedures considered for testing the marginal distribution of each topological feature (TF) of the robust test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
errorPlots(pd, nsim=nsim, tStat = FALSE, global=FALSE)

```


```{r, eval=FALSE, fig.width=8, fig.height=10, fig.cap="QQ-plot for the inference procedures considered for the marginal distribution of each topological feature (TF) of the parametric test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
qqPlots(pd, nsim=nsim, tStat = TRUE, global=FALSE)

```

```{r, eval=FALSE, fig.width=8, fig.height=10, fig.cap="QQ-plot for the inference procedures considered for the marginal distribution of each topological feature (TF) of the robust test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
qqPlots(pd, nsim=nsim, tStat = FALSE, global=FALSE)

```





### Global distributions

```{r, eval=FALSE, fig.width=8, fig.height=10, fig.cap="Actual versus target type 1 error rates for the inference procedures considered for testing the distribution of the global maximum of each topological feature (TF) of the parametric test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
errorPlots(pd, nsim=nsim, tStat = TRUE, global=TRUE)

```

```{r, eval=FALSE, fig.width=8, fig.height=10, fig.cap="Actual versus target type 1 error rates for the inference procedures considered for testing the distribution of the global maximum of each topological feature (TF) of the robust test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
errorPlots(pd, nsim=nsim, tStat = FALSE, global=TRUE)

```



```{r, eval=FALSE, fig.width=8, fig.height=10, fig.cap="QQ-plot for the inference procedures considered for the distribution of the global maximum of each topological feature (TF) of the parametric test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
qqPlots(pd, nsim=nsim, tStat = TRUE, global=TRUE)

```

```{r, eval=FALSE, fig.width=8, fig.height=10, fig.cap="QQ-plot for the inference procedures considered for the distribution of the global maximum of each topological feature (TF) of the robust test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
qqPlots(pd, nsim=nsim, tStat = FALSE, global=TRUE)

```




## Fake group variable

Group was simulated independently of the imaging data and the bootstrap samples of the imaging data were modeled and tested on 3 degrees of freedom.
Type 1 error rates and QQ-plot are given below.


### Marginal distribution 

```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="Actual versus target type 1 error rates for the inference procedures considered for testing the marginal distribution of each topological feature (TF) of the parametric test statistics image."}
pd = plotData(fakeGroupSimConfig$output, stats=c("Maxima", "Mass; cft = 0.01", "Mass; cft = 0.001", "Extent; cft = 0.01", "Extent; cft = 0.001", "pTFCE"))
nsim = fakeGroupSimConfig$nsim
errorPlots(pd, nsim=nsim, tStat = TRUE, global=FALSE)

```

```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="Actual versus target type 1 error rates for the inference procedures considered for testing the marginal distribution of each topological feature (TF) of the robust test statistics image."}
errorPlots(pd, nsim=nsim, tStat = FALSE, global=FALSE)

```


```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="QQ-plot for the inference procedures considered for the marginal distribution of each topological feature (TF) of the parametric test statistics image."}
qqPlots(pd, nsim=nsim, tStat = TRUE, global=FALSE)

```

```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="QQ-plot for the inference procedures considered for the marginal distribution of each topological feature (TF) of the robust test statistics image."}
qqPlots(pd, nsim=nsim, tStat = FALSE, global=FALSE)

```





### Global distributions

```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="Actual versus target type 1 error rates for the inference procedures considered for testing the distribution of the global maximum of each topological feature (TF) of the parametric test statistics image."}
errorPlots(pd, nsim=nsim, tStat = TRUE, global=TRUE)

```

```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="Actual versus target type 1 error rates for the inference procedures considered for testing the distribution of the global maximum of each topological feature (TF) of the robust test statistics image."}
errorPlots(pd, nsim=nsim, tStat = FALSE, global=TRUE)

```



```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="QQ-plot for the inference procedures considered for the distribution of the global maximum of each topological feature (TF) of the parametric test statistics image."}
qqPlots(pd, nsim=nsim, tStat = TRUE, global=TRUE)

```

```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="QQ-plot for the inference procedures considered for the distribution of the global maximum of each topological feature (TF) of the robust test statistics image."}
qqPlots(pd, nsim=nsim, tStat = FALSE, global=TRUE)

```








## Age continuous covariate fit with splines

The imaging data were residualized to age and other covariates using equation (13) in the main paper, then bootstrap samples of the residuals were modeled with age using splines on 4 degrees of freedom.
the test was for the nonlinear effect of age over the linear age effect on 3 degrees of freedom.


### Marginal distribution 

```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="Actual versus target type 1 error rates for the inference procedures considered for testing the marginal distribution of each topological feature (TF) of the parametric test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
pd = plotData(ageSplineSimConfig$output, stats=c("Maxima", "Mass; cft = 0.01", "Mass; cft = 0.001", "Extent; cft = 0.01", "Extent; cft = 0.001", "pTFCE"))
nsim = ageSplineSimConfig$nsim
errorPlots(pd, nsim=nsim, tStat = TRUE, global=FALSE)

```

```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="Actual versus target type 1 error rates for the inference procedures considered for testing the marginal distribution of each topological feature (TF) of the robust test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
errorPlots(pd, nsim=nsim, tStat = FALSE, global=FALSE)

```


```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="QQ-plot for the inference procedures considered for the marginal distribution of each topological feature (TF) of the parametric test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
qqPlots(pd, nsim=nsim, tStat = TRUE, global=FALSE)

```

```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="QQ-plot for the inference procedures considered for the marginal distribution of each topological feature (TF) of the robust test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
qqPlots(pd, nsim=nsim, tStat = FALSE, global=FALSE)

```





### Global distributions

```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="Actual versus target type 1 error rates for the inference procedures considered for testing the distribution of the global maximum of each topological feature (TF) of the parametric test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
errorPlots(pd, nsim=nsim, tStat = TRUE, global=TRUE)

```

```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="Actual versus target type 1 error rates for the inference procedures considered for testing the distribution of the global maximum of each topological feature (TF) of the robust test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
errorPlots(pd, nsim=nsim, tStat = FALSE, global=TRUE)

```



```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="QQ-plot for the inference procedures considered for the distribution of the global maximum of each topological feature (TF) of the parametric test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
qqPlots(pd, nsim=nsim, tStat = TRUE, global=TRUE)

```

```{r, eval=TRUE, fig.width=8, fig.height=10, fig.cap="QQ-plot for the inference procedures considered for the distribution of the global maximum of each topological feature (TF) of the robust test statistics image."}
#pdf('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_fakePolynomial.pdf')
qqPlots(pd, nsim=nsim, tStat = FALSE, global=TRUE)

```


```{r, eval=FALSE}
knitr::kable(cbind(pd$tfce$tStatmap_pbjNormT, pd$tfce$HC3RobustStatmap_pbjNormT[,2]), format = 'latex')
```

