---
title: "Documentation for running pbj simulations on AWS"
author: "Simon Vandekar"
date: "2/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=15, fig.height=9)
path = Sys.getenv('PATH')
path = Sys.setenv('PATH'=paste(path, '/home/rstudio/.local/bin', sep=':'))
```

## AWS machine image setup

I use the directions [here](https://jagg19.github.io/2019/08/aws-r/#short-easy) to create an AMI to run Rstudio on.
The `Welcome.R` script in the [NIsim](https://github.com/simonvandekar/NIsim) package has code to setup this machine image with Dropbox access to the files.


## Setup simulations

```{r simconfig}
# install the latest versions of the packages to perform these analyses.
devtools::install_github('simonvandekar/pbj', ref='ftest')
devtools::install_github('simonvandekar/NIsim')
### LIBRARIES ###
library(RNifti)
library(parallel)
library(splines)
library(mmand)
library(fslr)
library(progress)
library(abind)
library(pbj)
library(PDQutils)
library(NIsim)



### LOAD IN DATA FROM DROPBOX ###
dbimagedir = '~/Dropbox (VUMC)/pbj/data/abide/neuroimaging/cpac/alff_cropped'
dbresimagedir = '~/Dropbox (VUMC)/pbj/data/abide/neuroimaging/cpac/alff_res'
dbdatafile = '~/Dropbox (VUMC)/pbj/data/abide/demographic/n1035_phenotypic_20190509.rds'
maskfile = '~/Dropbox (VUMC)/pbj/data/abide/neuroimaging/cpac/n1035_mask.nii.gz'


# load in data and get directories
dat = readRDS(dbdatafile)
dat$imgname = paste(dat$file_id, 'alff.nii.gz', sep='_')
dat$files = file.path(dbimagedir, dat$imgname)




### SIMULATION PARAMETERS ###
fakePolySimConfig = list(
  # vector of sample sizes to simulate
  ns = c(50, 100, 200, 400),
  # number of simulations to run
  nsim=5,
  # number of bootstraps
  nboot = 500,
  # number of permutations
  nperm = 500,
  # cluster forming thresholds
  cfts.s = c(0.1, 0.25, 0.4),
  cfts.p = c(0.01, 0.001),
  
  # radius for spheres of signal.
  rs=c(8),
  
  #### MODEL FORMULAS FOR SIMULATIONS ####
  formres = as.formula( paste0(" ~ dx_group + sex + ns(func_mean_fd, df=10) + ns(age_at_scan, df=10)" )),
  # need age_at_scan in both models for testing nonlinear functions
  form = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan + fake_covariate1 + scale(fake_covariate1^2) + scale(fake_covariate1^3)" )),
  formred = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan + fake_covariate1")),
  #  weights for each subject. Can be a character vector
  W = c("func_mean_fd"),
  # where to put residuals
  resdir = dbresimagedir,
  # where to output results
  simdir = '~/temp',
  dat = dat,
  mask = maskfile,
  output = '~/Dropbox (VUMC)/pbj/pbj_ftest/df2_polynomial.rdata',
  ncores = 32
)
# use betas = 0 for global null
# parameters = betas * sd(y)/sd(x).
fakePolySimConfig$betas = rep(0, length(simConfig$rs))

### OTHER SIMULATION SETUPS ###
# 1 DF is real group
groupSimConfig = fakePolySimConfig
groupSimConfig$form = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan + dx_group" ))
groupSimConfig$form = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan" ))
groupSimConfig$output = '~/Dropbox (VUMC)/pbj/pbj_ftest/df1_dxgroup.rdata'

# FAKE POLY IS 2 DOF

# 3 DOF
fakeGroupSimConfig = fakePolySimConfig
fakeGroupSimConfig$form = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan + fake_group" ))
fakeGroupSimConfig$formred = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan" ))
fakeGroupSimConfig$output = '~/Dropbox (VUMC)/pbj/pbj_ftest/df3_fakegroup.rdata'

# 4 DOF motion
motionSplineSimConfig = fakePolySimConfig
motionSplineSimConfig$form = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan + ns(func_mean_fd, df=4)" ))
motionSplineSimConfig$formred = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan + func_mean_fd" ))
motionSplineSimConfig$output = '~/Dropbox (VUMC)/pbj/pbj_ftest/df4_motionspline.rdata'

# 5 DOF
ageSplineSimConfig = fakePolySimConfig
ageSplineSimConfig$form = as.formula(paste0(" ~ sex + func_mean_fd + ns(age_at_scan, df=5)" ))
ageSplineSimConfig$formred = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan" ))
motionSplineSimConfig$output = '~/Dropbox (VUMC)/pbj/pbj_ftest/df5_agespline.rdata'
```

# Simulation functions
```{r simulationFunctions}
# Function that gets observed and bootstrap values from a pbj object.
getBoots = function(pbjObj){
  cftnames = grep('cft', names(pbjObj), value=TRUE)
  out = do.call(cbind, lapply(pbjObj[ cftnames ], function(x) x$boots))
  colnames(out) = cftnames
  ccomps = lapply(pbjObj[ cftnames ], function(x) x$obs)
  return(list(obs=ccomps, boots=out))
}

# Statistic function to get objects for pbjInference
simStats = function(image, mask, thrs){
  c(maximum = max(c(image)), pbj::cluster(image, mask, thrs))
}

# simfunc should contain a data argument, which is defined within runSim
# Other arguments are identical across simulation runs.
simFunc = function(lmfull, lmred, mask, data, nboot, cfts){
  # generate fake covariates
  data$fake_group = factor(ceiling(ppoints(nrow(data))*4 ) )
  data$fake_covariate1 = rnorm(nrow(data))

  # compute Statistical maps
  # Residuals from null model to use for permutation procedures
  nullResids = lmPBJ(data$images, form=lmred, formred=~ 1, mask=mask, data=data, transform = 'none', robust=FALSE )$sqrtSigma
  nullResids = scale(t(nullResids), center=FALSE, scale=TRUE)

  # t transform, robust, estimate covariance
  tRobustStatmap = lmPBJ(data$images, form=lmfull, formred=lmred, mask=mask, data=data, transform = 't' )
  # t transform, classical, estimate covariance
  tStatmap = lmPBJ(data$images, form=lmfull, formred=lmred, mask=mask, data=data, transform = 't', robust=FALSE)
  # no transform, robust, estimate covariance
  robustStatmap = lmPBJ(data$images, form=lmfull, formred=lmred, mask=mask, data=data, transform ='none')
  # no transform, classical, estimate covariance
  plainStatmap = lmPBJ(data$images, form=lmfull, formred=lmred, mask=mask, data=data, transform = 'none', robust=FALSE)
  resids = plainStatmap$sqrtSigma

  statmaps = grep('Statmap', ls(), value=TRUE)
  out = list()
  # doesn't matter which statmap we use here
  thrs = (cfts^2*plainStatmap$rdf) + plainStatmap$df
  # Apply each of the sampling methods
  for(statmapname in statmaps){

    ### BOOTSTRAP METHODS
    statmap = get(statmapname)
    # normal bootstrap
    #pbjNorm = getBoots(pbjSEI(statmap, nboot = nboot, cfts.s = cfts))
    pbjNorm = pbjInference(statmap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats)
    pbjNormT = pbjInference(statmap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats, tboot=TRUE)
    # Rademacher bootstrap
    pbjRad = pbjInference(statmap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats, rboot = function(n){ 2*rbinom(n, size=1, prob=0.5)-1})
    pbjRadT = pbjInference(statmap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats, rboot = function(n){ 2*rbinom(n, size=1, prob=0.5)-1}, tboot=TRUE)
    #pbjRad = getBoots(pbjSEI(statmap, nboot = nboot, cfts.s = cfts, rboot = function(n){ 2*rbinom(n, size=1, prob=0.5)-1}))
    # t bootstrap
    #pbjNormT = getBoots(pbjSEI(statmap, nboot = nboot, cfts.s = cfts, tboot=TRUE))
    # Rademacher t
    #pbjRadT = getBoots(pbjSEI(statmap, nboot = nboot, cfts.s = cfts, rboot = function(n){ 2*rbinom(n, size=1, prob=0.5)-1}, tboot=TRUE))

    ### PERMUTATION METHODS
    # Q needs to absolute value of Q for permutation procedure. This multiplies by the sign of the residuals to do that.
    permMap = statmap
    # check dimensions
    permMap$sqrtSigma = sweep(permMap$sqrtSigma, MARGIN = c(1,2), STATS = sign(resids), FUN="*")
    # permutation # need to get null residuals from lmPBJ for this one
    perm = function(n){
      nullResids[sample(1:n),]
    }
    pbjPerm = pbjInference(permMap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats, rboot = perm)
    pbjPermT = pbjInference(permMap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats, rboot = perm, tboot=TRUE)

    # exchangeable permutation
    nullResids = sign(nullResids)
    exchPerm = function(n){
      nullResids[sample(1:n),]
    }
    pbjPermExch = pbjInference(permMap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats, rboot = exchPerm)
    pbjPermExchT = pbjInference(permMap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats, rboot = exchPerm, tboot=TRUE)

    # collect output
    PBJnames = grep('^pbj', ls(), value=TRUE)
    allnames = paste(statmapname, PBJnames, sep='_')
    out[allnames] = lapply(PBJnames, get, pos = environment())

    ### REPEAT ALL WITH INDEPENDENCE SPATIAL COVARIANCE ASSUMPTION
    # nonrobust methods won't be different, because covariance is same for all statistics.
  }
  return(out)
}

#debug(pbjInference)
#test = simFunc(simConfig$form, simConfig$formred, simConfig$mask, readRDS(file.path(simdirs$simdir[101], 'data.rds')), simConfig$nboot, simConfig$cfts.s)
```

# Run simulations
```{r runSims}
sims = grep('SimConfig', ls(), value=TRUE)
for(sim in sims[1]){
  # get simulation configuration for this simulation
  simConfig = get(sim) 
  ### SETUP THE SIMULATION ANALYSIS ###
  # subsets dataset to all people who have the variables
  simConfig$dat = simConfig$dat[apply(!is.na(simConfig$dat[ ,c(all.vars(as.formula(simConfig$formres)), simConfig$W)]), 1, all), ]
  # Create residualized images
  if(class(simConfig$formres)=='formula' | is.character(simConfig$formres)){
    simConfig$dat$rfiles = file.path(simConfig$resdir, basename(simConfig$dat$files))
    if(!all(file.exists(simConfig$dat$rfiles))){
      pbj::residualizeImages(files=simConfig$dat$files, dat=simConfig$dat, mask=simConfig$mask, form=simConfig$formres,
                             outfiles=simConfig$dat$rfiles, mc.cores=computeConfig$ncores)
    }
    simConfig$dat$files = simConfig$dat$rfiles
    # clean up. May not be necessary
    gc()
  }
  
  simdirs = simSetup(simConfig$dat$files, data=simConfig$dat, outdir=simConfig$simdir, nsim=simConfig$nsim, ns=simConfig$ns, mask=simConfig$mask, rs=simConfig$rs, betas=simConfig$betas )
  
  results = runSim(simdirs$simdir, method='synthetic',
       simfunc = simFunc, mask = simConfig$mask,
simfuncArgs = list(
  lmfull= simConfig$form,
  lmred = simConfig$formred,
  mask = simConfig$mask, nboot=simConfig$nboot, cfts=simConfig$cfts.s), ncores = computeConfig$ncores)

dir.create(dirname(simConfig$output), showWarnings = FALSE, recursive = TRUE)
# clean up files
save.image(file=simConfig$output)
unlink(list.files(tempdir(), full.names = TRUE))
gc()
unlink(simdirs)
}
```


# plotting results
```{r, eval=TRUE}
# plotting function for below sections
plots = function(rdata){
  load(rdata)
simdirs$results = results# lapply(results, simplify2array)
x =simdirs[simdirs$n==100,]
simdirs[, paste0('obsMaxClust_cft.s', simConfig$cfts.s)] = do.call(rbind, by(simdirs, simdirs$n, function(x) do.call(rbind, lapply(x$results, function(y) unlist(y[['obs']]) ) ) ))

cex=1.5
par(mgp=c(1.7,.7,0), lwd=1.5, lend=2, cex.lab=0.8*cex, cex.axis=0.8*cex, cex.main=1*cex, mfrow=c(1,1), mar=c(2.8,2.8,1.8,.2), bty='l')
layout(mat=matrix(1:(length(simConfig$cfts.s)*length(simConfig$ns)), nrow=length(simConfig$cfts.s)) )
# axes are based on tail quantiles
probs = seq(0.75, 1, length.out=100) #length.out=pmin(simConfig$nsim, simConfig$nboot)

trash = by(simdirs, simdirs$n, function(df){
  for(cftInd in 1:length(simConfig$cfts.s)){
    ylims = range(sapply(df$results, function(x) range(quantile(x$boot[[cftInd]][[1]], probs=probs))))
    colname = paste0('obsMaxClust_cft.s', simConfig$cfts.s[cftInd])
    x = df[,colname]
    xlims = range(quantile(x, probs=probs) )
    xaxlab = c(0.9, 0.95, 0.99, 0.999) 
    xaxt = quantile(x, probs=xaxlab)
    plot(x, ylim=ylims, xlim=xlims, type='n', xlab='Observed cluster quantile', ylab='Estimated cluster quantile', main=paste('n =', df$n[1], 'cft =', simConfig$cfts.s[cftInd]))
    #axis(side=1, at=xaxt, labels=xaxlab)
    abline(v=xaxt, col='orange', lty=2)
    for(ind in 1:simConfig$nsim){
      points(quantile(x, probs=probs), quantile(df$results[[ind]]$boot[[cftInd]][[1]], probs=probs), type='l')
    }
    abline(a=0,b=1, col='blue')
  }
})

trash = by(simdirs, simdirs$n, function(df){
  for(cftInd in 1:length(simConfig$cfts.s)){
    ylims = range(sapply(df$results, function(x) range(quantile(x$boot[[cftInd]][[1]], probs=probs))))
    colname = paste0('obsMaxClust_cft.s', simConfig$cfts.s[cftInd])
    x = df[,colname]
    xlims = range(quantile(x, probs=probs) )
    xaxlab = c(0.9, 0.95, 0.99, 0.999) 
    xaxt = quantile(x, probs=xaxlab)
    y=colMeans(do.call(rbind, lapply(1:nrow(df), function(ind) quantile(df$results[[ind]]$boot[[cftInd]][[1]], probs=xaxlab)<df[ind,colname])) )
    plot(1-xaxlab, y, type='b', xlab='Target type 1 error', ylab='Actual type 1 error', xlim=range(c(y, 1-xaxlab)), ylim=range(c(y, 1-xaxlab)), , main=paste('n =', df$n[1], 'cft =', simConfig$cfts.s[cftInd]))
    abline(a=0,b=1, col='blue')
  }
})
}
```

## Group covariate

```{r, eval=TRUE}
plots('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_group_covariate.rdata')
```



## Independent continuous covariates

```{r, eval=TRUE}
plots('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_independent_covariates.rdata')
```


## Polynomial continuous covariate

Testing the second and third degree terms of a polynomial covariate.

```{r, eval=TRUE}
plots('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_polynomial_covariate.rdata')
```
