---
title: "Documentation for running pbj simulations on AWS"
author: "Simon Vandekar"
date: "2/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval=FALSE, message=FALSE, warning=FALSE, fig.width=15, fig.height=9)
path = Sys.getenv('PATH')
path = Sys.setenv('PATH'=paste(path, '/home/rstudio/.local/bin', sep=':'))
```

## AWS machine image setup

I use the directions [here](https://jagg19.github.io/2019/08/aws-r/#short-easy) to create an AMI to run Rstudio on.
The `Welcome.R` script in the [NIsim](https://github.com/simonvandekar/NIsim) package has code to setup this machine image with Dropbox access to the files.


## Setup simulations

```{r simconfig}
# install the latest versions of the packages to perform these analyses.
devtools::install_github('simonvandekar/pbj', ref='ftest')
devtools::install_github('simonvandekar/NIsim')
### LIBRARIES ###
library(RNifti)
library(parallel)
library(splines)
library(mmand)
library(fslr)
library(progress)
library(abind)
library(pbj)
library(PDQutils)
library(NIsim)



### LOAD IN DATA FROM DROPBOX ###
dbimagedir = '~/Dropbox (VUMC)/pbj/data/abide/neuroimaging/cpac/alff'
dbresimagedir = '~/Dropbox (VUMC)/pbj/data/abide/neuroimaging/cpac/alff_res'
maskfile = '~/Dropbox (VUMC)/pbj/data/abide/neuroimaging/cpac/n1035_mask.nii.gz'
#dbimagedir = '~/Dropbox (VUMC)/pbj/data/abide/neuroimaging/cpac/alff_cropped/'
#dbresimagedir = '~/Dropbox (VUMC)/pbj/data/abide/neuroimaging/cpac/alff_cropped_res/'
#maskfile = '~/Dropbox (VUMC)/pbj/data/abide/neuroimaging/cpac/cropped_n1035_mask.nii.gz'
dbdatafile = '~/Dropbox (VUMC)/pbj/data/abide/demographic/n1035_phenotypic_20190509.rds'


# load in data and get directories
dat = readRDS(dbdatafile)
dat$imgname = paste(dat$file_id, 'alff.nii.gz', sep='_')
dat$files = file.path(dbimagedir, dat$imgname)




### SIMULATION PARAMETERS ###
fakePolySimConfig = list(
  # vector of sample sizes to simulate
  ns = c(200, 400),
  # number of simulations to run
  nsim=100,
  # number of bootstraps
  nboot = 500,
  # cluster forming thresholds
  cfts.s = c(0.1, 0.25, 0.4),
  cfts.p = c(0.01, 0.001),
  
  # radius for spheres of signal.
  rs=c(8),
  
  #### MODEL FORMULAS FOR SIMULATIONS ####
  formres = as.formula( paste0(" ~ dx_group + sex + ns(func_mean_fd, df=10) + ns(age_at_scan, df=10)" )),
  # need age_at_scan in both models for testing nonlinear functions
  form = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan + fake_covariate1 + scale(fake_covariate1^2) + scale(fake_covariate1^3)" )),
  formred = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan + fake_covariate1")),
  #  weights for each subject. Can be a character vector
  W = c("func_mean_fd"),
  # where to put residuals
  resdir = dbresimagedir,
  # where to output results
  simdir = '~/temp',
  dat = dat,
  mask = maskfile,
  output = '~/Dropbox (VUMC)/pbj/pbj_ftest/df2_polynomial.rdata',
  ncores = 64
)
# use betas = 0 for global null
# parameters = betas * sd(y)/sd(x).
fakePolySimConfig$betas = rep(0, length(fakePolySimConfig$rs))

### OTHER SIMULATION SETUPS ###
# 1 DF is real group
groupSimConfig = fakePolySimConfig
groupSimConfig$form = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan + dx_group" ))
groupSimConfig$formred = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan" ))
groupSimConfig$output = '~/Dropbox (VUMC)/pbj/pbj_ftest/df1_dxgroup.rdata'

# FAKE POLY IS 2 DOF

# 3 DOF
fakeGroupSimConfig = fakePolySimConfig
fakeGroupSimConfig$form = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan + fake_group" ))
fakeGroupSimConfig$formred = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan" ))
fakeGroupSimConfig$output = '~/Dropbox (VUMC)/pbj/pbj_ftest/df3_fakegroup.rdata'

# 4 DOF motion
motionSplineSimConfig = fakePolySimConfig
motionSplineSimConfig$form = as.formula(paste0(" ~ sex + age_at_scan + ns(func_mean_fd, df=5)" ))
motionSplineSimConfig$formred = as.formula(paste0(" ~ sex + age_at_scan + func_mean_fd" ))
motionSplineSimConfig$output = '~/Dropbox (VUMC)/pbj/pbj_ftest/df4_motionspline.rdata'

# 5 DOF
ageSplineSimConfig = fakePolySimConfig
ageSplineSimConfig$form = as.formula(paste0(" ~ sex + func_mean_fd + ns(age_at_scan, df=6)" ))
ageSplineSimConfig$formred = as.formula(paste0(" ~ sex + func_mean_fd + age_at_scan" ))
motionSplineSimConfig$output = '~/Dropbox (VUMC)/pbj/pbj_ftest/df5_agespline.rdata'
```

# Simulation functions
```{r simulationFunctions}
# Function that gets observed and bootstrap values from a pbj object.
getBoots = function(pbjObj){
  cftnames = grep('cft', names(pbjObj), value=TRUE)
  out = do.call(cbind, lapply(pbjObj[ cftnames ], function(x) x$boots))
  colnames(out) = cftnames
  ccomps = lapply(pbjObj[ cftnames ], function(x) x$obs)
  return(list(obs=ccomps, boots=out))
  
}

# Statistic function to get objects for pbjInference
simStats = function(image, mask, thrs){
  c(maximum = max(c(image)), pbj::cluster(image, mask, thrs))
}

# simfunc should contain a data argument, which is defined within runSim
# Other arguments are identical across simulation runs.
simFunc = function(lmfull, lmred, mask, data, nboot, cfts){
  # generate fake covariates
  data$fake_group = factor(ceiling(ppoints(nrow(data))*4 ) )
  data$fake_covariate1 = rnorm(nrow(data))

  # t transform, robust, estimate covariance
  tRobustStatmap = lmPBJ(data$images, form=lmfull, formred=lmred, mask=mask, data=data, transform = 't' )
  # t transform, classical, estimate covariance
  #tStatmap = lmPBJ(data$images, form=lmfull, formred=lmred, mask=mask, data=data, transform = 't', robust=FALSE)
  # Doesn't scale residuals by hat matrix diagonal
  #tPermStatmap = lmPBJ(data$images, form=lmfull, formred=lmred, mask=mask, data=data, transform = 't', HC3=FALSE )
  # no transform, robust, estimate covariance
  #robustStatmap = lmPBJ(data$images, form=lmfull, formred=lmred, mask=mask, data=data, transform ='none')
  # no transform, classical, estimate covariance
  #plainstatmap = lmPBJ(data$images, form=lmfull, formred=lmred, mask=mask, data=data, transform = 'none', robust=FALSE)

  statmaps = c('tRobustStatmap')
  out = list()
  # doesn't matter which statmap we use here
  thrs = (cfts^2*tRobustStatmap$rdf) + tRobustStatmap$df
  # Apply each of the sampling methods
  for(statmapname in statmaps){

    ### BOOTSTRAP METHODS
    statmap = get(statmapname)
    # normal bootstrap
    #pbjNorm = getBoots(pbjSEI(statmap, nboot = nboot, cfts.s = cfts))
    if(statmapname %in% c('tRobustStatmap')){
      pbjNormT = pbjInference(statmap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats, method='t')
      # Rademacher bootstrap
      pbjRadT = pbjInference(statmap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats, rboot = function(n){ 2*rbinom(n, size=1, prob=0.5)-1}, method='t')
    }
    # if(statmapname=='tRobustStatmap'){
    #   pbjRadRobust = pbjInference(statmap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats, rboot = function(n){ 2*rbinom(n, size=1, prob=0.5)-1}, method='robust')
    #   pbjNormRobust = pbjInference(statmap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats, method='robust')
    #   # exchangeable permutation
    #   # scale out sign of Q from sqrtSigma -- sign comes from "permExch" function
    #   statmap$sqrtSigma = sweep(statmap$sqrtSigma, MARGIN = c(1,2), STATS = sign(resids), FUN="*")
    #   pbjPermExch = pbjInference(statmap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats, rboot = exchPerm, method='robust')
    # }
    # if(statmapname == 'tStatmap'){
    #   pbjNorm = pbjInference(statmap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats, method='regular')
    #   pbjRad = pbjInference(statmap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats, rboot = function(n){ 2*rbinom(n, size=1, prob=0.5)-1}, method='regular')
    # }
    # 
    # if(statmapname=='tPermStatmap'){
    #   # now scale out residuals from sqrtSigma -- scale comes from "perm" function
    #   statmap$sqrtSigma = sweep(statmap$sqrtSigma, MARGIN = c(1,2), STATS = resids, FUN="/")
    #   pbjPerm = pbjInference(statmap, nboot = nboot, thr = thrs, mask=statmap$mask, statistic=simStats, rboot = perm, method='robust')
    # }
    # collect output
    PBJnames = grep('^pbj', ls(), value=TRUE)
    allnames = paste(statmapname, PBJnames, sep='_')
    out[allnames] = lapply(PBJnames, get, pos = environment())
    rm(PBJnames)

    ### REPEAT ALL WITH INDEPENDENCE SPATIAL COVARIANCE ASSUMPTION
    # nonrobust methods won't be different, because covariance is same for all statistics.
  }
  return(out)
}

#debug(pbjInference)
#simConfig = get("fakePolySimConfig")
#simdirs = simSetup(simConfig$dat$files, data=simConfig$dat, outdir=simConfig$simdir, nsim=simConfig$nsim, ns=simConfig$ns, mask=simConfig$mask, rs=simConfig$rs, betas=simConfig$betas )
#simtime = system.time(test <- simFunc(simConfig$form, simConfig$formred, simConfig$mask, readRDS(file.path(simdirs$simdir[200], 'data.rds')), simConfig$nboot, simConfig$cfts.s))
```

# Run simulations
```{r runSims}
sims = grep('SimConfig', ls(), value=TRUE)
for(sim in sims[3]){
  # get simulation configuration for this simulation
  simConfig = get(sim) 
  ### SETUP THE SIMULATION ANALYSIS ###
  # subsets dataset to all people who have the variables
  simConfig$dat = simConfig$dat[apply(!is.na(simConfig$dat[ ,c(all.vars(as.formula(simConfig$formres)), simConfig$W)]), 1, all), ]
  # Create residualized images
  if(class(simConfig$formres)=='formula' | is.character(simConfig$formres)){
    simConfig$dat$rfiles = file.path(simConfig$resdir, basename(simConfig$dat$files))
    if(!all(file.exists(simConfig$dat$rfiles))){
      pbj::residualizeImages(files=simConfig$dat$files, dat=simConfig$dat, mask=simConfig$mask, form=simConfig$formres,
                             outfiles=simConfig$dat$rfiles, mc.cores=simConfig$ncores)
    }
    simConfig$dat$files = simConfig$dat$rfiles
    # clean up. May not be necessary
    gc()
  }
  
  simdirs = simSetup(simConfig$dat$files, data=simConfig$dat, outdir=simConfig$simdir, nsim=simConfig$nsim, ns=simConfig$ns, mask=simConfig$mask, rs=simConfig$rs, betas=simConfig$betas )
  
  
  #time = system.time(test <- simFunc(simConfig$form, simConfig$formred, simConfig$mask, readRDS(file.path(simdirs$simdir[10], 'data.rds')), simConfig$nboot, simConfig$cfts.s) )
  
  # mix this up so that large sample simulations aren't all dropped on one "thread".
  simdirs = simdirs[sample(1:nrow(simdirs)),]
  results = runSim(simdirs$simdir, method='synthetic',
                   simfunc = simFunc, mask = simConfig$mask,
                   simfuncArgs = list(
                     lmfull= simConfig$form,
                     lmred = simConfig$formred,
                     mask = simConfig$mask, nboot=simConfig$nboot, cfts=simConfig$cfts.s), ncores = simConfig$ncores)
  
  dir.create(dirname(simConfig$output), showWarnings = FALSE, recursive = TRUE)
  # clean up files
  save.image(file=simConfig$output)
  unlink(list.files(tempdir(), full.names = TRUE))
  gc()
  unlink(simdirs)
}
# shuts down aws machine
Sys.sleep(60*5)
system('sudo shutdown -h now')
```


# plotting results
```{r, eval=TRUE}
# for each method plot:
# qqplot of maximum value for each sample size
# qqplot of max cluster size for each cft and sample size
# plotting function for below sections
plots = function(rdata){
  load(rdata)
  simdirs$results = results# lapply(results, simplify2array)
  methods = names(simdirs$results[[1]])
  stats = c("Max", paste('cft =', simConfig$cfts.s) )
  
  # graphical parameters
  cex=1.5
  par(mgp=c(1.7,.7,0), lwd=1.5, lend=2, cex.lab=0.8*cex, cex.axis=0.8*cex, cex.main=1*cex, mfrow=c(1,1), mar=c(2.8,2.8,1.8,.2), bty='l', oma=c(0,0,2,0))
  layout(mat=matrix(1:(length(stats)*length(simConfig$ns)), nrow=length(stats)) )
  # axes are based on tail quantiles
  probs = seq(0.75, 1, length.out=100)
  
  for(method in methods){
  obsStat = do.call(rbind, lapply(simdirs$results, function(y) if(is.null(y)) NA else sapply(y[[method]][['obsStat']],function(z) {suppressWarnings(res <- max(z)); res[is.infinite(res)] = 0; res} ) ) )
  # These colnames were sample size dependent
  simdirs[, stats] = obsStat
  
  simdirs$boots = lapply(simdirs$results, function(y) do.call(rbind, lapply(y[[method]][['boots']], function(z0) sapply(z0, function(z) {suppressWarnings(res <- max(z)); res[is.infinite(res)] = 0; res} ) ) ) )
 #length.out=pmin(simConfig$nsim, simConfig$nboot)
  
  trash = by(simdirs, simdirs$n, function(df){
    for(cftInd in 1:length(stats)){
      ylims = range(sapply(df$boots, function(x) range(quantile(x[,cftInd], probs=probs, na.rm=TRUE) ) ), na.rm=TRUE )
      colname = stats[cftInd]
      x = df[,colname ]
      xlims = range(quantile(x, probs=probs, na.rm=TRUE) )
      xaxlab = c(0.9, 0.95, 0.99, 0.999) 
      xaxt = quantile(x, probs=xaxlab, na.rm = TRUE)
      plot(x, ylim=ylims, xlim=xlims, type='n', xlab='Observed quantile', ylab='Estimated quantile', main=paste('n =', df$n[1],  colname))
      #axis(side=1, at=xaxt, labels=xaxlab)
      abline(v=xaxt, col='orange', lty=2)
      for(ind in 1:simConfig$nsim){
        if(!is.null(df$boots[[ind]])) points(quantile(x, probs=probs, na.rm=TRUE), quantile(df$boots[[ind]][,cftInd], probs=probs, na.rm=TRUE), type='l')
      }
      abline(a=0,b=1, col='blue')
    }
  })
      mtext(method, outer=TRUE)
  
  trash = by(simdirs, simdirs$n, function(df){
    for(cftInd in 1:length(stats)){
      ylims = range(sapply(df$boots, function(x) range(quantile(x[,cftInd], probs=probs, na.rm = TRUE))), na.rm=TRUE)
      x = df[,stats[cftInd] ]
      xlims = range(quantile(x, probs=probs, na.rm=TRUE))
      xaxlab = c(0.9, 0.95, 0.99, 0.999) 
      xaxt = quantile(x, probs=xaxlab, na.rm=TRUE)
      y=colMeans(do.call(rbind, lapply(1:nrow(df), function(ind) quantile(df$boots[[ind]][,cftInd], probs=xaxlab, na.rm =TRUE)<df[ind,stats[cftInd] ]) ), na.rm=TRUE )
      plot(1-xaxlab, y, type='b', xlab='Target type 1 error', ylab='Actual type 1 error', xlim=range(c(y, 1-xaxlab)), ylim=range(c(y, 1-xaxlab)), main=paste('n =', df$n[1],  stats[cftInd]) )
      abline(a=0,b=1, col='blue')
    }
  })
   mtext(method, outer=TRUE)
  }
}
```

## Group covariate




## Independent continuous covariates


## Polynomial continuous covariate

Testing the second and third degree terms of a polynomial covariate.

```{r, eval=TRUE}
#debug(plots)
plots('~/Dropbox (VUMC)/pbj/pbj_ftest/df2_polynomial.rdata')
```
